---
title: "Regression exercise: Water salinity"
author: "Constance Delannoy"
output: html_document
---
<!-- # Purpose: training on logistic regression using CalCOFI dataset and prompt from kagle -->
<!-- # Author: Constance -->
<!-- # Date started: 4/17/2018 -->

Dataset: CalCOFI; Over 60 years of oceanographic data

Questions: Is there a relationship between water salinity & water temperature? Can you predict the water temperature based on salinity?

Method: Linear regression

Findings: Temperature is not a good predictor of salinity except when paired with Depth and O2 concentration.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)

# Libraries
library(tidyverse)
library(ggplot2)
library(corrplot)
library(gridExtra)
library(stargazer)
library(mgcv)
library(jtools)
library(lubridate)
library(ggmap)
library(zoo)

# directories
calcofi_dir <- "C:/Users/cdelannoy/Documents/Projects/courses/bayes/kaggle/lm_calcofi/"
input_dir <- file.path(calcofi_dir, "input")
output_dir <- file.path(calcofi_dir, "output")
```

```{r loading data and preliminary processing}
calcofi_raw <- read_csv("N:\\Transfer\\CDelannoy\\calcofi\\input\\bottle.csv")
calcofi_cast<- read_csv("N:\\Transfer\\CDelannoy\\calcofi\\input\\cast.csv", col_types = cols(Cruz_Num = col_character(), .default = col_guess()))

# Getting rid of "quality code" and "precision" variables, irrelevant variables (and replicate variables??) to only keep actual measures
calcofi_df <- calcofi_raw %>% 
  select(-ends_with("Aq"),
         -ends_with("1q"),
         -ends_with("2q"),
         -ends_with("3q"),
         -ends_with("4q"),
         -ends_with("Ap"),
         -ends_with("1p"),
         -ends_with("2p"),
         -ends_with("3p"),
         -ends_with("4p"),
         -ends_with("qual"),
         -ends_with("prec"),
         -ends_with("tq"),
         -ends_with("qua"),
         -ends_with("qu"),
         -`DIC Quality Comment`,
         -R_SAMP,
         -Cst_Cnt,
         -Btl_Cnt,
         -Depth_ID,
         -Sta_ID,
         -starts_with("DIC"),
         -starts_with("TA"),
         -starts_with("MeanA"),
         -IncTim,
         -starts_with("C14A") #,
         #-R_SALINITY
         ) %>% 
 mutate_if(is.character, as.numeric) %>% 
  filter(R_SALINITY > 5)

```

```{r cast cleaning}
calcofi_cast_clean <- calcofi_cast %>% 
  mutate_at(vars(Date), mdy) %>% 
  mutate(month_year = paste0(Month, "-", Year)) %>% 
  mutate_at(vars(month_year), as.yearmon, "%m-%Y")

```

```{r}
merged_df <- left_join(calcofi_raw, calcofi_cast_clean, by = "Cst_Cnt")

# quantile depth grouping
plot_data <- merged_df %>% mutate(Depth_bucket = case_when(
  Depthm == 0 ~ "zero",
  Depthm < 46 & Depthm > 0 ~ "Q1",
  Depthm < 125 & Depthm >= 46 ~ "Q2",
  Depthm < 226.8 & Depthm >= 125 ~ "Q3",
  Depthm < 300 & Depthm >= 226.8 ~ "Q4",
  Depthm >= 300 ~ "max", 
  TRUE ~ NA_character_)
  ) %>%
  mutate_at(vars(Depth_bucket), as.factor)


temp_date_p <- ggplot(plot_data,
                      aes(x = Month, y = T_degC, group = Depth_bucket, color = Depth_bucket)) +
              geom_point()
temp_date_p
```

```{r}

```



```{r samples over time}
time_data <- calcofi_cast_clean %>% 
  group_by(Date) %>% 
  summarise(n_samples = n())

time_plot <- ggplot(time_data, aes(x = Date, y = n_samples)) +
  geom_line() +
  ggtitle("number of samples over time")

time_plot
```

```{r samples per month}
samples_month <- calcofi_cast_clean %>% 
  group_by(Month) %>% 
  summarise(n_samples = n())

month_plot <- ggplot(samples_month, aes(x = Month, y = n_samples)) +
  geom_line() +
  ggtitle("Number of samples per month") +
  scale_x_discrete(limits = as.character(unique(samples_month$Month) %>% order())) +
  ylab("Number of samples")

month_plot
```

```{r}
samples_ymonth <- calcofi_cast_clean %>% 
  group_by(month_year) %>% 
  summarise(n_samples = n())

ymonth_plot <- ggplot(samples_ymonth, aes(x = month_year, y = n_samples)) +
  geom_line() +
  ggtitle("Number of samples per month and year") +
  ylab("Number of samples")

ymonth_plot
```


```{r location of samples}
mapWorld <- ggplot() +
  borders("world", colour = "gray50", fill = "gray50") +
  geom_point(data = calcofi_cast_clean, aes(x = Lon_Deg, y = Lat_Deg), colour = "red", size = 3) +
  ggtitle("Location of water samples")

mapWorld
```
```{r}


westcoastmap <- ggplot() +
  borders("world", colour = "gray50", fill = "gray50") +
  geom_point(data = plot_data %>% filter(Depth_bucket == "zero",
                                         Month == 1),
             aes(x = Lon_Deg, y = Lat_Deg, colour = T_degC), size = 2) +
  ggtitle("Location of water samples") +
  scale_x_continuous( limits = c( min(plot_data$Lon_Deg) - 1, max(plot_data$Lon_Deg) + 1 ) , expand = c( 0, 0) ) + 
  scale_y_continuous( limits = c( min(plot_data$Lat_Deg) - 1, max(plot_data$Lat_Deg) +1 ) , expand = c( 0 , 0 ) ) 

westcoastmap

unique(merged_df$Month)
```


# {.tabset}

## Exploration and data cleaning {.tabset}


### Outcome variable

<!-- The outcome variable (temperature in degress Celsius) looks reasonably normal, so we do not feel the need to transform it. -->

<!-- ```{r exploration} -->
<!-- outcome_plot <- ggplot(calcofi_raw, aes(T_degC)) + -->
<!--   geom_histogram() + -->
<!--   ggtitle("distribution of outcome variable (temperature in C)") -->
<!-- outcome_plot -->

<!-- ``` -->

```{r}
outcome_na <- sum(is.na(calcofi_df$T_degC))
calcofi_nona <- calcofi_raw %>% 
  filter(!is.na(T_degC)) %>% 
  filter(R_SALINITY > 4.57)
```

The number of NA values in the outcome variable is `r outcome_na`. We filter the dataset to only keep rows without NAs.

```{r outcome var plot}

ggplot(data = NULL, aes(x = calcofi_nona$T_degC)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("Temperature") +
  ylab("density") +
  ggtitle("Distribution of Temperature")
```


We now turn to the independent variables.

### Independent variables

The two main variables of interest seem to have outliers in both directions.
```{r}
main_dep_plot <- ggplot(calcofi_nona, aes(Salnty)) +
  geom_histogram() +
  ggtitle("distribution of main dependent variable of interest (salinity of water)")
main_dep_plot
```


```{r}
whisker_salnty <- ggplot(calcofi_nona, aes(x = "", y = Salnty)) +
  geom_boxplot()           

whisker_salnty
```

We look at the second measure of salinity. This variable has one outlier at R_SALINITY = 4.57. We delete this value to get the distribution below.

```{r}
main_dep_plot2 <- ggplot(calcofi_nona, aes(R_SALINITY)) +
  geom_histogram() +
  ggtitle("distribution of main dependent variable of interest (reported salinity)")
main_dep_plot2
```

Before removing the variable from our data, we confirm that the two variables are correlated (which they are). 
Because the correlation between the two variables is 1, we only keep one of them, Salnty.

```{r}
salnty_nona <- calcofi_nona %>% 
  filter(!is.na(R_SALINITY))
cat("The correlation between R_SALINITY and Salnty is", round(cor(salnty_nona$R_SALINITY, salnty_nona$Salnty), 4))
```


We remove variables related to replication bottles because they are highly correlated with one another and because there are very few non-NA values.
```{r correlations for replicate variables}
correlated_candidates <- calcofi_raw %>% 
  select(starts_with("DIC"),
         starts_with("TA"),
         starts_with("MeanA"),
         starts_with("C14A")) %>% 
  na.omit() %>% 
  mutate_all(as.numeric)

#corrplot(cor(correlated_candidates))
```


As a next step, we check the amount of NAs in all of our columns. Some of them have more than 50% NA values (note we have already filtered out NA rows on our outcome variable). We remove these variables from our dataset.

```{r NA check, results="asis"}
na_table <- calcofi_df %>% 
  summarise_all(funs(sum(is.na(.))/n())) %>% 
  gather(variable, sum_na) %>% 
  filter(sum_na > 0.5)

cols_to_remove <- na_table$variable

stargazer(na_table,
          type = "html",
          summary = FALSE)
```

```{r calcofi df ready for analysis}
calcofi_corrtrim <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>% 
  na.omit()
```



Finally, we check the correlations of other variables in our dataset.
We remove the following variables:

* R_TEMP because it is measuring the same thing as T_degC
* R_Depth because it is measuring the same thing as Depthm
* R_O2, and R_PRES because they are both highly correlated with Depthm
* R_POTEMP, R_SIGMA, R_SVA, DYNHT because they are highly correlated with T_degC
* STheta, Oxy_umol/kg, Reclnd R_O2, R_O2Sat and O2Sat because they are highly correlated ith O2ml_L

```{r total corr check}
corrplot(cor(calcofi_corrtrim), 
         type = "upper", 
         diag = TRUE)

calcofi_reg <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>%  
  select(-R_TEMP,
         -R_Depth,
         -R_O2,
         -R_PRES,
         -R_POTEMP,
         -R_SIGMA,
         -R_SVA,
         -R_DYNHT,
         -STheta,
        # -`Oxy_µmol/Kg`,
         -RecInd,
         -R_O2,
         -R_O2Sat,
         -O2Sat,
         -R_SALINITY) 

calcofi_reg <- calcofi_reg[, -5] %>% # removing `Oxy_µmol/Kg` column
  na.omit()

stopifnot(!("Oxy" %in% colnames(calcofi_reg)))

```


The resulting dataset has 4 columns and `r nrow(calcofi_reg)` rows.

Removing all NAs in the remaining rows leaves us with `r nrow(calcofi_reg)` rows out of `r nrow(calcofi_raw)`, or `r round(nrow(calcofi_reg)/nrow(calcofi_raw), 2)` % of the data.

```{r}
corrplot(cor(calcofi_reg), 
         type = "upper", 
         diag = TRUE)

```

___Heatmaps___

```{r}
heatmap1 <- ggplot(calcofi_reg, aes(x = Depthm, y = Salnty)) +
  geom_tile(aes(fill = T_degC)) +
  ggtitle("Heatmap of Depth vs Salinity")

heatmap1
```
```{r}
heatmap2 <- ggplot(calcofi_reg, aes(x = O2ml_L, y = Salnty)) +
  geom_tile(aes(fill = T_degC)) +
  ggtitle("Heatmap of O2 concentration vs Salinity")

heatmap2
```

```{r}
heatmap3 <- ggplot(calcofi_reg, aes(x = O2ml_L, y = Depthm)) +
  geom_tile(aes(fill = T_degC)) +
  ggtitle("Heatmap of O2 concentration vs Depth")

heatmap3
```


___Data sections___

```{r create cut-off variable}
calcofi_reg <- calcofi_reg %>% 
  mutate(sections = factor(case_when(Depthm <= 250 & O2ml_L >= 6                               ~ 1,
                                     Depthm <= 20 & O2ml_L < 6 & O2ml_L >= 5.5                 ~ 2,
                                     Depthm > 20 & Depthm <= 250 & O2ml_L < 6 & O2ml_L >= 5.5  ~ 3,
                                     Depthm <= 250 & O2ml_L < 5.5 & O2ml_L >= 5                ~ 4,
                                     Depthm <= 250 & O2ml_L < 5 & O2ml_L >= 4                  ~ 5,
                                     Depthm <= 250 & O2ml_L < 4 & O2ml_L >= 3                  ~ 6,
                                     Depthm <= 250 & O2ml_L < 3 & O2ml_L >= 2                  ~ 7,
                                     Depthm <= 250 & O2ml_L < 2 & O2ml_L >= 1.5                ~ 8,
                                     Depthm <= 250 & O2ml_L < 1.5                              ~ 9,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 6 & O2ml_L >= 2   ~ 10,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 2 & O2ml_L >= 1.5 ~ 11,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 1.5               ~ 12,
                                     Depthm > 500 & Depthm < 1500                              ~ 13,
                                     Depthm >= 1500                                            ~ 14)))


```


First, we take a look at the distribution of O2 vs Depth to pick the cut-offs.
```{r} 
ggplot(calcofi_reg, aes(x = Depthm, y = O2ml_L)) +
  geom_point(aes(color = sections)) +
  ggtitle("Cut-off for values of Depth for next graphs")
```

Key for what each section number represents:

| Key | Depthm | O2 concentration|
|---|---|---|
| 1  | less than 250m     | greater than 6ml/L|
| 2  | less than 20m      | btw 5.5 and 6ml/L|
| 3  | btw 20 and 250m    | btw 5.5 and 6ml/L|
| 4  | less than 250m     | btw 5 and 5.5ml/L|
| 5  | less than 250m     | btw 4 and 5ml/L|
| 6  | less than 250m     | btw 3 and 4 ml/L|
| 7  | less than 250m     | btw 2 and 3 ml/L|
| 8  | less than 250m     | btw 1.5 and 2 ml/L|
| 9  | less than 250m     | less than 1.5ml/L|
| 10  | btw 250 and 500m  | btw 2 and 6 ml/L|
| 11 | btw 250 and 500m   | btw 1.5 and 2 ml/L|
| 12 | btw 250 and 500m   | less than 1.5ml/L|
| 13 | btw 500 and 1500m  | any |
| 14 | greater than 1500m | any |

We then check the amount of data in each section

```{r}
N_per_section <- calcofi_reg %>% 
  group_by(sections) %>% 
  summarise(sample = n())

knitr::kable(N_per_section, row.names = FALSE)
```


We decide to create sections of the data based on the above graph.
We now show the distribution of O2 vs Depth for each section.
```{r}
ggplot(calcofi_reg, aes(y = T_degC, x = Salnty)) +
  geom_point() +
  facet_wrap(~sections, ncol = 5) +
  ggtitle("Temperature vs Salinity for various values of O2 concentration and Depth") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


___PCA___

We conduct principal component analysis (PCA) using prcomp on the O2 and Depth variables.

```{r}
pca <- prcomp(calcofi_reg %>% select(O2ml_L, Depthm), scale. = TRUE)
biplot(pca, scale=0)
```
The principal components loadings are below:
```{r}
pca$rotation
```
Here is the standard deviation and variance of each component:

```{r}
pca_st_dev <- pca$sdev
pca_var <- pca$sdev^2
```
* Standard Deviation: `r pca_st_dev`
* Variance: `r pca_var`

And the proportion of the variance explained by the components:
```{r}
prop_varex <- pca_var/sum(pca_var)
```
* First component: `r prop_varex[1]`
* Second component: `r prop_varex[2]`

Based on this result, we select the first component for our model.
We now plot Salinity vs PCA1.

```{r}
pca_col <- data.frame(PCA1 = pca$x[,1])
ggplot(bind_cols(calcofi_reg, pca_col), aes(x = PCA1, y = Salnty, color = T_degC)) +
  geom_point() +
  ggtitle("Salinity vs PCA1 with Temperature as color gradient") +
  xlim(-4, 1)
```


## Modeling {.tabset}


```{r model fitting, results="asis"}
gam_lm <- gam(T_degC ~ s(Depthm) + O2ml_L + Salnty, data = calcofi_reg)

gam_lm_int <- gam(T_degC ~ te(Depthm, Salnty, O2ml_L), data = calcofi_reg)
```

Let's compare both GAM models on AIC.

```{r}
AIC(gam_lm, gam_lm_int)
```


### Model with GAM

The following resource was helpful in understanding GAM: https://people.maths.bris.ac.uk/~sw15190/mgcv/tampere/mgcv.pdf 

based on the scatter plot below, it still seems that there is an effect we are not quite capturing, based on the negative diagonal trend lines we can make out.

```{r residuals vs fitted gam}
ggplot(NULL, aes(x = gam_lm$fitted.values, y = gam_lm$residuals)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Residuals vs Fitted values for regression with gam") +
  xlab("fitted values") +
  ylab("residuals")

```

This plot shows that GAM did a good job addressing the heteroscedasticity issue we have been having. This plot looks the best so far, with an almost even distribution of variance for any predicted values.

```{r residuals vs x gam}
ggplot(NULL, aes(x = gam_lm$fitted.values, y = calcofi_reg$T_degC)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Observed vs predicted values for regression with gam") +
  xlab("predicted values (Temperature in Celsius)") +
  ylab("observed values (Temperature in Celsius)")

```

<!-- ```{r plots round 3 gam} -->
<!-- # dep <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Depthm))+ -->
<!-- #   geom_point() + -->
<!-- #   xlab("Depth (meters)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # temp <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Salnty))+ -->
<!-- #   geom_point() + -->
<!-- #   xlab("Salinity (standard scale)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # o2 <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$O2ml_L))+ -->
<!-- #   geom_point()+ -->
<!-- #   xlab("O2 concentration (ml/L)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # grid.arrange(dep, temp, o2, ncol = 2) -->
<!-- ``` -->

The residual plot below shows a normal distribution of errors. 

```{r errors gam}
ggplot(data = NULL, aes(x = gam_lm$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("count") +
  ggtitle("Distribution of residuals")

```

```{r GAM splines, warning = FALSE}
par(mfrow=c(1,3))
splines_3d <- vis.gam(gam_lm, ticktype = "detailed", se = 2, theta= -40)
splines_3d2 <- vis.gam(gam_lm, theta= 40)
splines_3d3 <- vis.gam(gam_lm, plot.type = "contour")
```

```{r more GAM plots}
salnty_plot <- ggplot(data = calcofi_reg, aes(x = Salnty)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightblue", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Salnty")

depthm_plot <- ggplot(data = calcofi_reg, aes(x = Depthm)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightpink", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

o2_plot <- ggplot(data = calcofi_reg, aes(x = O2ml_L)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightgreen", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

grid.arrange(salnty_plot, depthm_plot, o2_plot, ncol = 3)
```

```{r more GAM}
par(mfrow=c(1,2))
splines <- plot(gam_lm, pers = TRUE)
plot(gam_lm, residuals = TRUE, pch = 19, main = "residual plot for spline term (Depth)")
```

#### Looking at some of the pre-defined summary plots from mgcv package and additional deviance plots
```{r}
par(mfrow=c(2,3))
gam.check(gam_lm)
plot(fitted(gam_lm), residuals(gam_lm), main = "fitted values vs residuals")
plot(calcofi_reg$T_degC, residuals(gam_lm), main = "observed values for Temperature vs residuals")
```


### Model with GAM and interaction terms

We start with some summary statistics for the model.

```{r}
# sse <- sum(gam_lm_int$fitted.values - gam_lm_int$y)^2
# ssr <- sum(gam_lm_int$fitted.values - mean(gam_lm_int$y))^2 
# sst <- sse + ssr
# 
# r_2 <- 1 - sse/sst
# adj_r_2 <- 1 - sse*length(gam_lm_int$y - 1)/(sst*(length(gam_lm_int$y) - length(gam_lm_int$coefficients)))
# 
# knitr::kable(gam_lm_int$coefficients, col.names = c("Coefficient"))
summary(gam_lm_int)
```


We now plot the residuals for this model.

```{r errors gam with interactions}
ggplot(data = NULL, aes(x = gam_lm_int$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("count") +
  ggtitle("Distribution of residuals")

```


```{r GAM splines 2, warning = FALSE}
par(mfrow=c(1,3))
splines_3d <- vis.gam(gam_lm_int, ticktype = "detailed", se = 2, theta= -40)
splines_3d2 <- vis.gam(gam_lm_int, theta= 40)
splines_3d3 <- vis.gam(gam_lm_int, plot.type = "contour")
```

```{r more GAM plots 2}
salnty_plot <- ggplot(data = calcofi_reg, aes(x = Salnty)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightblue", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Salnty")

depthm_plot <- ggplot(data = calcofi_reg, aes(x = Depthm)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightpink", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

o2_plot <- ggplot(data = calcofi_reg, aes(x = O2ml_L)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightgreen", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

grid.arrange(salnty_plot, depthm_plot, o2_plot, ncol = 3)
```

```{r more GAM 2}
# par(mfrow=c(1,2))
# splines <- plot(gam_lm_int)#, pers = TRUE)
# plot(gam_lm_int, residuals = TRUE, pch = 19, main = "residual plot for spline term (Depth)")
```


```{r summary plots GAM 2}
par(mfrow=c(2,3))
gam.check(gam_lm_int)
plot(fitted(gam_lm_int), residuals(gam_lm_int), main = "fitted values vs residuals")
plot(calcofi_reg$T_degC, residuals(gam_lm_int), main = "observed values for Temperature vs residuals")
```

The predictions are in the range we expect, although sub-zero water temperatures seem a bit odd:

```{r}
summary(gam_lm_int$fitted.values)
```


Now let's pot some interactions!
High values of depth produce extreme predicted temperatures... we produce plots as-is, and with those extreme values removed.

```{r interaction lines GAM}
inter_plot <- function(df, factor_var, gam_model, x_var, extreme = TRUE){
  var <- rlang::sym(factor_var)
  plot_dat <- data.frame(df %>% select(-!!var) %>% {.[1:100,]}) %>% 
      mutate(!!var := rep(quantile(df[[factor_var]], probs = c(0, 0.25, 0.5, 0.75, 1)), times = 20)) %>% 
    select(-T_degC)
   
  pred <- predict(gam_model, newdata = plot_dat, se.fit = TRUE) 
  plot_dat <- transform(plot_dat,
                        fitted = pred$fit,
                        se     = pred$se.fit,
                        fac   = factor(plot_dat[[factor_var]]))
  
  if(extreme == FALSE){
    plot_dat <- plot_dat %>% 
      filter(fitted < 50,
             fitted > -50)
  }
  
  crit <- qt(0.975, df = df.residual(gam_model))
  
  plot_dat <- transform(plot_dat,
                        upper = fitted + (crit * se),
                        lower = fitted - (crit * se))

  inter_gam_plot <- ggplot(plot_dat, aes_string(x_var, y = "fitted", group = "fac")) +
      geom_line(aes_string(colour = "fac")) +   # draw trend lines
      geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
      theme_bw() +                        # minimal theme
      theme(legend.position = "none") +   # no legend
      labs(y = expression(Temperature ~ (degree*C)), x = NULL) +
      facet_wrap(~ fac, ncol = 5) +    # facet on month
      ggtitle(paste("Predictions of temperature vs", x_var, "based on quantile of", factor_var))  
    
  inter_gam_plot
}

inter_plot(calcofi_reg, "Salnty", gam_lm_int, "Depthm")
```
```{r}
inter_plot(calcofi_reg, "Depthm", gam_lm_int, "Salnty")
```



```{r interaction lines GAM Depth}
inter_plot(calcofi_reg, "O2ml_L", gam_lm_int, "Salnty")
```

Here are plots without extreme values

```{r}
inter_plot(calcofi_reg, "Salnty", gam_lm_int, "Depthm", extreme = FALSE)
```

```{r}
inter_plot(calcofi_reg, "Depthm", gam_lm_int, "Salnty", extreme = FALSE)
```

```{r interaction lines GAM Depth 2}
inter_plot(calcofi_reg, "O2ml_L", gam_lm_int, "Salnty", extreme = FALSE)
```
```{r}
plots <- pairs(calcofi_reg)
```

```{r}
# fit1 <- gam(T_degC ~ te(Depthm, Salnty, O2ml_L), data = calcofi_reg)
fit1 <- gam(T_degC ~ te(Depthm, Salnty), data = calcofi_reg)
plot.gam(fit1)
```

```{r}
fit2 <- gam(T_degC ~ te(Salnty, O2ml_L), data = calcofi_reg)
plot(fit2)
```

```{r}
fit3 <- gam(T_degC ~ te(Depthm, O2ml_L), data = calcofi_reg)
plot(fit3)
```
