---
title: "Regression exercise: Water salinity"
author: "Constance Delannoy"
date: "April 19, 2018"
output: html_document
---
<!-- # Purpose: training on logistic regression using CalCOFI dataset and prompt from kagle -->
<!-- # Author: Constance -->
<!-- # Date started: 4/17/2018 -->

Dataset: CalCOFI; Over 60 years of oceanographic data

Questions: Is there a relationship between water salinity & water temperature? Can you predict the water temperature based on salinity?

Method: Linear regression

Findings: Temperature is not a good predictor of salinity except when paired with Depth and O2 concentration.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)

# Libraries
library(tidyverse)
library(ggplot2)
library(corrplot)
library(gridExtra)
library(stargazer)
library(mgcv)
library(jtools)

# directories
calcofi_dir <- "C:/Users/cdelannoy/Documents/Projects/courses/bayes/kaggle/lm_calcofi/"
input_dir <- file.path(calcofi_dir, "input")
output_dir <- file.path(calcofi_dir, "output")
```

```{r loading data and preliminary processing}
calcofi_raw <- read_csv(file.path(input_dir, "bottle.csv"))

# Getting rid of "quality code" and "precision" variables, irrelevant variables (and replicate variables??) to only keep actual measures
calcofi_df <- calcofi_raw %>% 
  select(-ends_with("Aq"),
         -ends_with("1q"),
         -ends_with("2q"),
         -ends_with("3q"),
         -ends_with("4q"),
         -ends_with("Ap"),
         -ends_with("1p"),
         -ends_with("2p"),
         -ends_with("3p"),
         -ends_with("4p"),
         -ends_with("qual"),
         -ends_with("prec"),
         -ends_with("tq"),
         -ends_with("qua"),
         -ends_with("qu"),
         -`DIC Quality Comment`,
         -R_SAMP,
         -Cst_Cnt,
         -Btl_Cnt,
         -Depth_ID,
         -Sta_ID,
         -starts_with("DIC"),
         -starts_with("TA"),
         -starts_with("MeanA"),
         -IncTim,
         -starts_with("C14A") #,
         #-R_SALINITY
         ) %>% 
 mutate_if(is.character, as.numeric) %>% 
  filter(R_SALINITY > 5)

```
# {.tabset}

## Exploration and data cleaning {.tabset}


### Outcome variable

<!-- The outcome variable (temperature in degress Celsius) looks reasonably normal, so we do not feel the need to transform it. -->

<!-- ```{r exploration} -->
<!-- outcome_plot <- ggplot(calcofi_raw, aes(T_degC)) + -->
<!--   geom_histogram() + -->
<!--   ggtitle("distribution of outcome variable (temperature in C)") -->
<!-- outcome_plot -->

<!-- ``` -->

```{r}
outcome_na <- sum(is.na(calcofi_df$T_degC))
calcofi_nona <- calcofi_raw %>% 
  filter(!is.na(T_degC)) %>% 
  filter(R_SALINITY > 4.57)
```

The number of NA values in the outcome variable is `r outcome_na`. We filter the dataset to only keep rows without NAs.

```{r outcome var plot}

ggplot(data = NULL, aes(x = calcofi_nona$T_degC)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("Temperature") +
  ylab("density") +
  ggtitle("Distribution of Temperature")
```


We now turn to the independent variables.

### Independent variables

The two main variables of interest seem to have outliers in both directions.
```{r}
main_dep_plot <- ggplot(calcofi_nona, aes(Salnty)) +
  geom_histogram() +
  ggtitle("distribution of main dependent variable of interest (salinity of water)")
main_dep_plot
```


```{r}
whisker_salnty <- ggplot(calcofi_nona, aes(x = "", y = Salnty)) +
  geom_boxplot()           

whisker_salnty
```

We look at the second measure of salinity. This variable has one outlier at R_SALINITY = 4.57. We delete this value to get the distribution below.

```{r}
main_dep_plot2 <- ggplot(calcofi_nona, aes(R_SALINITY)) +
  geom_histogram() +
  ggtitle("distribution of main dependent variable of interest (reported salinity)")
main_dep_plot2
```

Before removing the variable from our data, we confirm that the two variables are correlated (which they are). 
Because the correlation between the two variables is 1, we only keep one of them, Salnty.

```{r}
salnty_nona <- calcofi_nona %>% 
  filter(!is.na(R_SALINITY))
cat("The correlation between R_SALINITY and Salnty is", round(cor(salnty_nona$R_SALINITY, salnty_nona$Salnty), 4))
```


We remove variables related to replication bottles because they are highly correlated with one another and because there are very few non-NA values.
```{r correlations for replicate variables}
correlated_candidates <- calcofi_raw %>% 
  select(starts_with("DIC"),
         starts_with("TA"),
         starts_with("MeanA"),
         starts_with("C14A")) %>% 
  na.omit() %>% 
  mutate_all(as.numeric)

#corrplot(cor(correlated_candidates))
```


As a next step, we check the amount of NAs in all of our columns. Some of them have more than 50% NA values (note we have already filtered out NA rows on our outcome variable). We remove these variables from our dataset.

```{r NA check, results="asis"}
na_table <- calcofi_df %>% 
  summarise_all(funs(sum(is.na(.))/n())) %>% 
  gather(variable, sum_na) %>% 
  filter(sum_na > 0.5)

cols_to_remove <- na_table$variable

stargazer(na_table,
          type = "html",
          summary = FALSE)
```

```{r calcofi df ready for analysis}
calcofi_corrtrim <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>% 
  na.omit()
```



Finally, we check the correlations of other variables in our dataset.
We remove the following variables:

* R_TEMP because it is measuring the same thing as T_degC
* R_Depth because it is measuring the same thing as Depthm
* R_O2, and R_PRES because they are both highly correlated with Depthm
* R_POTEMP, R_SIGMA, R_SVA, DYNHT because they are highly correlated with T_degC
* STheta, Oxy_umol/kg, Reclnd R_O2, R_O2Sat and O2Sat because they are highly correlated ith O2ml_L

```{r total corr check}
corrplot(cor(calcofi_corrtrim), 
         type = "upper", 
         diag = TRUE)

calcofi_reg <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>%  
  select(-R_TEMP,
         -R_Depth,
         -R_O2,
         -R_PRES,
         -R_POTEMP,
         -R_SIGMA,
         -R_SVA,
         -R_DYNHT,
         -STheta,
        # -`Oxy_µmol/Kg`,
         -RecInd,
         -R_O2,
         -R_O2Sat,
         -O2Sat,
         -R_SALINITY) 

calcofi_reg <- calcofi_reg[, -5] %>% # removing `Oxy_µmol/Kg` column
  na.omit()

stopifnot(!("Oxy" %in% colnames(calcofi_reg)))

```


The resulting dataset has 4 columns and `r nrow(calcofi_reg)` rows.

Removing all NAs in the remaining rows leaves us with `r nrow(calcofi_reg)` rows out of `r nrow(calcofi_raw)`, or `r round(nrow(calcofi_reg)/nrow(calcofi_raw), 2)` % of the data.

```{r}
corrplot(cor(calcofi_reg), 
         type = "upper", 
         diag = TRUE)

```


## Modeling {.tabset}

*Note: I could not get a summary of the gam model with splines on each interaction term...

```{r model fitting, results="asis"}
calcofi_lm_Salnty <- lm(T_degC ~ ., dat = calcofi_reg)

calcofi_log <- calcofi_reg %>% 
  bind_cols(data.frame("log_Depthm" = log(calcofi_reg$Depthm))) %>% 
  filter(!is.infinite(log_Depthm)) %>% 
  select(-Depthm)
  
calcofi_lm_log <- lm(T_degC ~ ., dat = calcofi_log)
calcofi_lm_temp <- lm(T_degC ~ Salnty, dat = calcofi_reg)

interaction_model <- lm(T_degC ~ log_Depthm + O2ml_L + Salnty + log_Depthm*Salnty + O2ml_L*Salnty + O2ml_L*log_Depthm + log_Depthm*Salnty*O2ml_L, dat = calcofi_log)

gam_lm <- gam(T_degC ~ s(Depthm) + O2ml_L + Salnty, data = calcofi_reg)

gam_lm_int <- gam(T_degC ~ s(Depthm) + O2ml_L + Salnty + s(Depthm,Salnty) + O2ml_L*Salnty + s(Depthm,O2ml_L) + s(Depthm,Salnty,O2ml_L), data = calcofi_reg)

stargazer(calcofi_lm_temp,
          calcofi_lm_Salnty,
          calcofi_lm_log,
          interaction_model,
          gam_lm,
          type = "html",
          column.labels = c("lm with Salnty only", "Standard lm", "lm with log(depth)", "lm with s(Depthm)"),
          column.separate = c(1, 1, 2, 1))
```


### Model with salinity only

There is clearly a trend that our model is not catching here.

```{r residuals temp}
res_temp_plot <- ggplot(NULL, aes(x = calcofi_lm_temp$fitted.values, y = calcofi_lm_temp$residuals)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Residuals vs Fitted values for regression on salinity only") +
  xlab("fitted values") +
  ylab("residuals")

res_temp_plot
```

Our model does not seem very accurate, based on the plot below.

```{r residuals vs x for 1-pred model}
res_vs_y_plot_salinity <- ggplot(NULL, aes(x = calcofi_lm_temp$fitted.values, y = calcofi_reg$T_degC)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Observed vs predicted values for salinity-only model") +
  xlab("predicted values (Temperature in Celsius)") +
  ylab("observed values (Temperature in Celsius)")

res_vs_y_plot_salinity
```

The residuals are slightly right-skewed.

```{r errors salinity only}
ggplot(data = NULL, aes(x = calcofi_lm_temp$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("density") +
  ggtitle("Distribution of residuals")

```


### Standard model with Salnty

There seems to be a trend in the residuals, jusging by the two diagonal lines spanning the scatter plot below.

```{r residuals vs fitted}
res_plot <- ggplot(NULL, aes(x = calcofi_lm_Salnty$fitted.values, y = calcofi_lm_Salnty$residuals)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Residuals vs Fitted values for standard regression with Salnty") +
  xlab("fitted values") +
  ylab("residuals")

res_plot
```


This plot shows potential heteroscedasticity issues, as lower values of Temperatures have a smaller variance than high ones.

```{r residuals vs x}
res_vs_y_plot <- ggplot(NULL, aes(x = calcofi_lm_Salnty$fitted.values, y = calcofi_reg$T_degC)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Observed vs predicted values for standard regression with Salnty") +
  xlab("predicted values (Temperature in Celsius)") +
  ylab("observed values (Temperature in Celsius)")

res_vs_y_plot
```

To further investigate, we plot the outcome variable on each dependent variable.

It seems that salinity values are not a linear function of Depth. In a separate model, we take the log of this variable to try to smooth out its distribution. In a separate model, we use the gmcv package to transform this variable.

In addition, the salinity variable shows much heteroscedasticity, as small values of salinity have around a 10 degree difference in temperature, while high values of salinity have around 30 degrees of difference in temperature.

```{r plots round 3}
dep <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Depthm))+
  geom_point() +
  xlab("Depth (meters)") +
  ylab("Temperature (Celsius)")

temp <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Salnty))+
  geom_point() +
  xlab("Salinity (standard scale)") +
  ylab("Temperature (Celsius)")

o2 <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$O2ml_L))+
  geom_point()+
  xlab("O2 concentration (ml/L)") +
  ylab("Temperature (Celsius)")

grid.arrange(dep, temp, o2, ncol = 2)
```

The residuals follow a standard normal distribution.

```{r errors standard}
ggplot(data = NULL, aes(x = calcofi_lm_Salnty$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("density") +
  ggtitle("Distribution of residuals")

```

### Model with log(Depth)

The residuals of this model are the most randomly scattered so far.

```{r residuals log}
res_log_plot <- ggplot(NULL, aes(x = calcofi_lm_log$fitted.values, y = calcofi_lm_log$residuals)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Residuals vs Fitted values for model with log(Depth)") +
  xlab("fitted values") +
  ylab("residuals")

res_log_plot
```

This graph still shows a heterscedasticity problem, although less pronounced than when Depth is not logged.

```{r residuals vs x for log model}
ggplot(NULL, aes(x = calcofi_lm_log$fitted.values, y = calcofi_log$T_degC)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Observed vs predicted values for log model") +
  xlab("predicted values (Temperature in Celsius)") +
  ylab("observed values (Temperature in Celsius)")

```

The graph below confirms our variance issue is still there, as high values of log(Depth) have much lower variance than low ones.

```{r plot log}
ggplot(data = NULL, aes(y = calcofi_log$T_degC, x = calcofi_log$log_Depthm))+
  geom_point() +
  xlab("log(Depth)") +
  ylab("Temperature (Celsius)") +
  ggtitle("Distribution of Temperature vs log(Depth")

```

The errors have a standard normal distribution.

```{r errors log Depthm}
ggplot(data = NULL, aes(x = calcofi_lm_log$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("desnity") +
  ggtitle("Distribution of residuals")

```

### Interaction model with log(Depthm)

We used log_Depthm as interaction variable because:

* the basic lm using log_Depthm has a higher adjusted R squared
* it seems to have two "levels" based on the Temp vs log(Depthm) plot we've seen on the previous tab

Below are plots of the variable vs the two other independent variables, Salnty and o2_ml. Neither of those plots looks very good: the plot with Salnty looks like it has a heteroscedasticity problem, and the plot with o2_ml does not look linear.

```{r interaction exploration}
dep_sal <- ggplot(data = NULL, aes(x = calcofi_log$log_Depthm, y = calcofi_log$Salnty))+
  geom_point() +
  xlab("log depth (meters)") +
  ylab("Salinity (standard scale)")

dep_o2 <- ggplot(data = NULL, aes(x = calcofi_log$log_Depthm, y = calcofi_log$O2ml_L))+
  geom_point() +
  ylab("O2 concentration (ml/L)") +
  xlab("log depth (meters)")

grid.arrange(dep_sal, dep_o2, ncol = 2)

```

We now use a plot to understand the effect of the interaction terms between log(Depth), Salnty and O2ml_L on the model. Based on this plot, it seems like interaction term would be useful in making more precise predictions.

```{r interactive interaction}
interact_plot1 <- interact_plot(interaction_model, pred = "Salnty", modx = "log_Depthm", interval = TRUE, int.width = 0.2) +
  ggtitle("Plot of outcome vs Salnty \n based on value of log_Depthm")

interact_plot2 <- interact_plot(interaction_model, pred = "O2ml_L", modx = "log_Depthm", interval = TRUE) +
  ggtitle("Plot of outcome vs O2ml_L \n based on value of log_Depthm")

grid.arrange(interact_plot1, interact_plot2, ncol = 2)
```

We now check the distribution of errors, which looks normal. 

```{r errors interaction model}
ggplot(data = NULL, aes(x = interaction_model$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("density") +
  ggtitle("Distribution of residuals")

```


### Model with GAM

The following resource was helpful in understanding GAM: https://people.maths.bris.ac.uk/~sw15190/mgcv/tampere/mgcv.pdf 

based on the scatter plot below, it still seems that there is an effect we are not quite capturing, based on the negative diagonal trend lines we can make out.

```{r residuals vs fitted gam}
ggplot(NULL, aes(x = gam_lm$fitted.values, y = gam_lm$residuals)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Residuals vs Fitted values for regression with gam") +
  xlab("fitted values") +
  ylab("residuals")

```

This plot shows that GAM did a good job addressing the heteroscedasticity issue we have been having. This plot looks the best so far, with an almost even distribution of variance for any predicted values.

```{r residuals vs x gam}
ggplot(NULL, aes(x = gam_lm$fitted.values, y = calcofi_reg$T_degC)) +
  geom_point() +
  geom_smooth(method = lm,
              se = FALSE) +
  ggtitle("Observed vs predicted values for regression with gam") +
  xlab("predicted values (Temperature in Celsius)") +
  ylab("observed values (Temperature in Celsius)")

```

<!-- ```{r plots round 3 gam} -->
<!-- # dep <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Depthm))+ -->
<!-- #   geom_point() + -->
<!-- #   xlab("Depth (meters)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # temp <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$Salnty))+ -->
<!-- #   geom_point() + -->
<!-- #   xlab("Salinity (standard scale)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # o2 <- ggplot(data = NULL, aes(y = calcofi_reg$T_degC, x = calcofi_reg$O2ml_L))+ -->
<!-- #   geom_point()+ -->
<!-- #   xlab("O2 concentration (ml/L)") + -->
<!-- #   ylab("Temperature (Celsius)") -->
<!-- #  -->
<!-- # grid.arrange(dep, temp, o2, ncol = 2) -->
<!-- ``` -->

The residual plot below shows a normal distribution of errors. 

```{r errors gam}
ggplot(data = NULL, aes(x = gam_lm$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("count") +
  ggtitle("Distribution of residuals")

```

```{r GAM splines, warning = FALSE}
par(mfrow=c(1,3))
splines_3d <- vis.gam(gam_lm, ticktype = "detailed", se = 2, theta= -40)
splines_3d2 <- vis.gam(gam_lm, theta= 40)
splines_3d3 <- vis.gam(gam_lm, plot.type = "contour")
```

```{r more GAM plots}
salnty_plot <- ggplot(data = calcofi_reg, aes(x = Salnty)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightblue", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Salnty")

depthm_plot <- ggplot(data = calcofi_reg, aes(x = Depthm)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightpink", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

o2_plot <- ggplot(data = calcofi_reg, aes(x = O2ml_L)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm)), color = "lightgreen", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

grid.arrange(salnty_plot, depthm_plot, o2_plot, ncol = 3)
```

```{r more GAM}
par(mfrow=c(1,2))
splines <- plot(gam_lm, pers = TRUE)
plot(gam_lm, residuals = TRUE, pch = 19, main = "residual plot for spline term (Depth)")
```

#### Looking at some of the pre-defined summary plots from mgcv package and additional deviance plots
```{r}
par(mfrow=c(2,3))
gam.check(gam_lm)
plot(fitted(gam_lm), residuals(gam_lm), main = "fitted values vs residuals")
plot(calcofi_reg$T_degC, residuals(gam_lm), main = "observed values for Temperature vs residuals")
```


### Model with GAM and interaction terms

We start with some summary statistics for the model.

```{r echo = TRUE}
# sse <- sum(gam_lm_int$fitted.values - gam_lm_int$y)^2
# ssr <- sum(gam_lm_int$fitted.values - mean(gam_lm_int$y))^2 
# sst <- sse + ssr
# 
# r_2 <- 1 - sse/sst
# adj_r_2 <- 1 - sse*length(gam_lm_int$y - 1)/(sst*(length(gam_lm_int$y) - length(gam_lm_int$coefficients)))
# 
# knitr::kable(gam_lm_int$coefficients, col.names = c("Coefficient"))
summary(gam_lm_int)
```

Let's compare both GAM models on AIC.

```{r}
AIC(gam_lm, gam_lm_int)
```


<!-- |metric|value|formula| -->
<!-- |:----:|:---:|:-----:| -->
<!-- |sse   |`r sse`|`sum(gam_lm_int$fitted.values - gam_lm_int$y)^2`| -->
<!-- |ssr   |`r ssr`|`sum(gam_lm_int$fitted.values - mean(gam_lm_int$y))^2`| -->
<!-- |sst   |`r sst`|`sse + ssr`| -->
<!-- |R2    |`r r_2`|`1 - sse/sst`| -->
<!-- |adj R2|`r adj_r_2`|`1 - sse*length(gam_lm_int$y - 1)/(sst*(length(gam_lm_int$y) - length(gam_lm_int$coefficients)))`| -->

We now plot the residuals for this model.

```{r errors gam with interactions}
ggplot(data = NULL, aes(x = gam_lm_int$residuals)) +
  geom_histogram(aes(y = ..density..), 
                 color = "blue") +
  geom_density(alpha = 0.2, fill = "#FF6666") +
  xlab("residuals") +
  ylab("count") +
  ggtitle("Distribution of residuals")

```


```{r GAM splines 2, warning = FALSE}
par(mfrow=c(1,3))
splines_3d <- vis.gam(gam_lm_int, ticktype = "detailed", se = 2, theta= -40)
splines_3d2 <- vis.gam(gam_lm_int, theta= 40)
splines_3d3 <- vis.gam(gam_lm_int, plot.type = "contour")
```

```{r more GAM plots 2}
salnty_plot <- ggplot(data = calcofi_reg, aes(x = Salnty)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightblue", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Salnty")

depthm_plot <- ggplot(data = calcofi_reg, aes(x = Depthm)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightpink", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

o2_plot <- ggplot(data = calcofi_reg, aes(x = O2ml_L)) +
  geom_point(aes(y = T_degC)) +
  geom_line(aes(y = predict(gam_lm_int)), color = "lightgreen", alpha = 0.8) +
  ggtitle("comparison of predicted values of Temperature and actual values of \n temperature vs Depth")

grid.arrange(salnty_plot, depthm_plot, o2_plot, ncol = 3)
```

```{r more GAM 2}
par(mfrow=c(1,2))
splines <- plot(gam_lm_int, pers = TRUE)
plot(gam_lm_int, residuals = TRUE, pch = 19, main = "residual plot for spline term (Depth)")
```


```{r summary plots GAM 2}
par(mfrow=c(2,3))
gam.check(gam_lm_int)
plot(fitted(gam_lm_int), residuals(gam_lm_int), main = "fitted values vs residuals")
plot(calcofi_reg$T_degC, residuals(gam_lm_int), main = "observed values for Temperature vs residuals")
```

Now let's pot some interactions!

```{r interaction lines GAM}
inter_plot <- function(df, factor_var, gam_model, x_var){
  var <- rlang::sym(factor_var)

plot_dat <- data.frame(df %>% select(-!!var) %>% {.[1:100,]}) %>% 
    mutate(!!var := rep(quantile(df[[factor_var]], probs = c(0, 0.25, 0.5, 0.75, 1)), times = 20))
  
pred <- predict(gam_model, newdata = plot_dat, se.fit = TRUE) 
plot_dat <- transform(plot_dat,
                      fitted = pred$fit,
                      se     = pred$se.fit,
                      fac   = factor(plot_dat[[factor_var]]))

crit <- qt(0.975, df = df.residual(gam_model))

plot_dat <- transform(plot_dat,
                      upper = fitted + (crit * se),
                      lower = fitted - (crit * se))


inter_gam_plot <- ggplot(plot_dat, aes_string(x_var, y = "fitted", group = "fac")) +
    geom_line(aes_string(colour = "fac")) +   # draw trend lines
    theme_bw() +                        # minimal theme
    theme(legend.position = "none") +   # no legend
    labs(y = expression(Temperature ~ (degree*C)), x = NULL) +
    facet_wrap(~ fac, ncol = 5) +    # facet on month
    ggtitle(paste("Predictions of temperature vs", x_var, "based on quantile of", factor_var))  
  
inter_gam_plot
}

inter_plot(calcofi_reg, "Salnty", gam_lm_int, "Depthm")
```
```{r}
inter_plot(calcofi_reg, "Depthm", gam_lm_int, "Salnty")
```



```{r interaction lines GAM Depth}
inter_plot(calcofi_reg, "O2ml_L", gam_lm_int, "Salnty")
```