---
title: "Why and how to use GAM"
subtitle: "with a focus on interactions"
author: "presented by Constance"
output:
  revealjs::revealjs_presentation
---

<style type="text/css">
h1.title {
  color: DarkBlue;
}
h1.subtitle {
  font-size: 38px;
  color: DarkBlue;
}
h1.author {
  font-size: 40px
  color: DarkBlue;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
```


## Why GAM?


* Principle of linear regression: linearity of outcome vs predictor

* Sometimes, that is not the case, so need to transform a predictor

* Ex:

## 
```{r linear kink ex, echo=FALSE}
kink_plot <- ggplot(NULL) +
  geom_line(aes(x = 0:5, y = 0:5))+
  geom_line(aes(x = 5:10, y = 5:0)) +
  geom_line(aes(x = 0:5, y = 10:5), linetype = 2) +
  geom_line(aes(x = 5:10, y = 5:10), linetype = 2) +
  theme(axis.title.y=element_blank(),
        axis.title.x = element_blank()) +
  ggtitle("Example of line to model using splines")

kink_plot
```


## MGCV package

* linear approximation by decomposition (into linear functions with different bases)

* `s()` acts as a factor variable, to be used to smooth unique variables by linear approximation

* `te()` produces a full tensor product smooth (useful for interactions)

* `ti()` produces a tensor product interaction (useful for interactions)

## Tensor productions te() vs ti()

* Used to smooth interaction terms

* Predictors need not be on the same scale

* `te()` incorporates the interaction and main effects: $f(x)+g(z)+h(x,z)$

* `ti()` incorporates only interaction, without main effects: $h(x,z)$

## 2-way interactions

* `y ~ s(x, by = z)` models linear interaction (if effect of $x$ on $z$ is linear)

* `y ~ te(x, z)` models non-linear interaction

* Note that when you use `by`, you are not getting main effects for that variable

## 3-way interactions

* The previous method can be extrapolated to 3-way (or more) interactions with `te()`

* `y ~ te(x, z, by = t)` models non-linear interaction between $x$ and $z$, with linear interaction with $t$

* `y ~ te(x, z, t)` models non-linear interaction between all three

## Sources

* https://stat.ethz.ch/R-manual/R-devel/library/mgcv/html/te.html

* https://stats.stackexchange.com/questions/86766/interactions-in-gam

* https://stats.stackexchange.com/questions/45446/intuition-behind-tensor-product-interactions-in-gams-mgcv-package-in-r 