---
title:
output: 
  revealjs::revealjs_presentation:
    theme: moon
    mathjax: local
    includes:
    self_contained: false
    reveal_plugins: ["notes", "menu"]
    reveal_options:
      slideNumber: true
      progress: true
      history: true
      overview: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# standard libraries
library(tidyverse)
library(rpart)
library(gridExtra)

# BART libraries
library(bartMachine)
library(BayesTree)
library(dbarts)
library(bcf)
```

```{r set-up, from bcf help page, include = FALSE}
# data generating process
p = 3 #two control variables and one moderator
n = 250
#
set.seed(1)

x = matrix(rnorm(n*p), nrow=n)

# create targeted selection
q = -1*(x[,1]>(x[,2])) + 1*(x[,1]<(x[,2]))

# generate treatment variable
pi = pnorm(q)
z = rbinom(n,1,pi)

# tau is the true (homogeneous) treatment effect
tau = (0.5*(x[,3] > -3/4) + 0.25*(x[,3] > 0) + 0.25*(x[,3]>3/4))

# generate the response using q, tau and z
mu = (q + tau*z)

# set the noise level relative to the expected mean function of Y
sigma = diff(range(q + tau*pi))/8

# draw the response variable with additive error
y = mu + sigma*rnorm(n)

# If you didn't know pi, you would estimate it here
pihat = pnorm(q)

# adding z nd pihat to input data for non-bcf models
x_nonbcf = cbind(x, z, pihat)
```


```{r bcf fit (as in BCF help page), include = FALSE}
bcf_fit   = bcf(y, z, x, x, pihat, nburn=2000, nsim=2000)
bt_fit    = BayesTree::bart(x_nonbcf, y, ntree = 200, ndpost = 2000, nskip = 2000)
bm_fit    = bartMachine(as.data.frame(x_nonbcf), y, num_trees = 200, num_burn_in = 2000, num_iterations_after_burn_in = 2000)
dbart_fit = dbarts::bart(x_nonbcf, y, ntree = 200, ndpost = 2000, nskip = 2000)
```


```{r functions}
plot_tree <- function(bart_fit, title, data){
  tree_data <- data.frame(data, yhat = colMeans(bart_fit$yhat))
  tree_fit <- rpart(yhat ~ ., data = tree_data)
  par(mar = c(0.2, 0.2, 0.7, 0.2))
  plot(tree_fit, uniform = TRUE, margin = 0.1, main = title)
  text(tree_fit, all = FALSE, use.n = TRUE)
}


plot_het_effects <- function(bart_fit, title, col_num = 2, data){
  bart_means <- map_dbl(1:ncol(bart_fit$yhat), function(c){mean(bart_fit$yhat[,c])})
  bart_lower <- map_dbl(1:ncol(bart_fit$yhat), function(c){quantile(bart_fit$yhat[,c], 0.025)})
  bart_upper <- map_dbl(1:ncol(bart_fit$yhat), function(c){quantile(bart_fit$yhat[,c], 0.875)})
  
  bart_true_est_means <- data.frame(means = c(bart_means, mu), groups = c(rep("estimate", length(bart_means)), rep("true", length(mu))), x = rep(data, 2))
  
  ggplot(NULL, aes(x=data[,col_num])) +
    geom_linerange(aes(ymax=bart_upper, ymin=bart_lower), color = "deepskyblue1", alpha = 0.5) +
    geom_point(aes(x = bart_true_est_means$x, y = bart_true_est_means$means, group = bart_true_est_means$groups, color = bart_true_est_means$groups), alpha = 0.5) +
    xlab(paste0("X", col_num)) +
    ylab("Posterior means of obs-specific impacts") +
    ggtitle(paste0(title)) +
    scale_color_manual("Groups", values=c("deepskyblue1",  "black"))
}

plot_sub_effects <- function(bart_fit, limit, model, col_num = 2, data){
    # browser()
  # Step 1: identify groups
  prediction <- t(bart_fit$yhat) %>% 
    as.data.frame()
  colnames(prediction) <- 1:2000

  prediction_X <- prediction %>% 
  bind_cols(data.frame(X = data[,col_num]))

  X_lower <- prediction_X %>% 
    filter(X < limit)
  
  X_upper <- prediction_X %>% 
    filter(X >= limit)
  
  # Step 2: take draw means
  X_lower_means <- colMeans(X_lower %>% select(-X))
  X_upper_means <- colMeans(X_upper %>% select(-X))
  X_means <- data.frame(means = c(X_lower_means, X_upper_means), group = c(rep("lower", length(X_lower_means)), rep("upper", length(X_upper_means)))) %>% 
    mutate_at(vars(group), as.factor)
  
  
  # Step 3: take difference in means for each draw
  X_means_diff <- X_upper_means - X_lower_means
  
  # Step 4: plot histogram of differences
  diff_plot <- ggplot(data = NULL) +
                geom_histogram(aes(X_means_diff)) +
                 ggtitle(paste0("Histogram of posterior draws of difference in group-specific impacts \n for", model)) +
                geom_vline(xintercept = quantile(X_means_diff, 0.025), color = "red") +
                geom_vline(xintercept = quantile(X_means_diff, 0.975), color = "red") +
                geom_vline(xintercept = mean(X_means_diff), color = "blue") +
                xlab("Difference in means")
    
  # Step 5: plot histogram of yhats above and below limits
  plot_breaks_norounding <- sort(c(min(X_lower_means), max(X_lower_means), min(X_upper_means), max(X_upper_means), mean(X_lower_means), mean(X_upper_means)))
  plot_breaks <- map_dbl(plot_breaks_norounding, round, 2)
  
  yhats_plot <- ggplot(data = X_means) +
                 geom_histogram(aes(means, group = group, fill = group), alpha = 0.5) +
                 ggtitle(paste0("Histogram of posterior draws of group-specific impacts \n for", model)) +
                 geom_vline(xintercept = quantile(X_lower_means, 0.025), color = "red", linetype = "dotted") +
                 geom_vline(xintercept = quantile(X_lower_means, 0.975), color = "red", linetype = "dotted") +
                 geom_vline(xintercept = mean(X_lower_means), color = "blue") +
                 geom_vline(xintercept = quantile(X_upper_means, 0.025), color = "red", linetype = "dotted") +
                 geom_vline(xintercept = quantile(X_upper_means, 0.975), color = "red", linetype = "dotted") +
                 geom_vline(xintercept = mean(X_upper_means), color = "blue") +
                 scale_x_continuous(breaks = plot_breaks) +
                 theme(axis.text.x = element_text(angle = 45, hjust = 1))
                 xlab("Means")
  
  return(grid.arrange(diff_plot, yhats_plot, ncol = 2))

}


```


# BCF 

## Tree

```{r bcf tree}
bcf_fit_edit <- bcf_fit
bcf_fit_edit$yhat <- bcf_fit_edit$tau 

plot_tree(bcf_fit_edit, "Tree from BCF model", data = x)
```
Based on tree, we choose to investigate X3.



## Heterogeneous effects

```{r bcf}
plot_het_effects(bcf_fit_edit, "Comparison of estimate vs actual heterogeneous effects with BCF", col_num = 3, data = x)
```

## Differences in subgroup treatment effects

Based on the tree, we decide to look at the differences between X3 smaller or greater than -0.3994.

```{r bcf difference in group-specific impacts}
plot_sub_effects(bcf_fit, -0.3994, "BCF model", col_num = 3, data = x)
```


# BayesTree 

## Tree

```{r bt tree}
bt_fit$yhat <- bt_fit$yhat.train
plot_tree(bt_fit, "Tree from BayesTree model", data = x_nonbcf)
```
Based on tree, we choose to investigate pihat.



## Heterogeneous effects

```{r bt}
plot_het_effects(bt_fit, "Comparison of estimate vs actual heterogeneous effects with BayesTree", col_num = 5, data = x_nonbcf)
```

## Differences in subgroup treatment effects

Based on the tree, we decide to look at the differences between pihat smaller or greater than 0.5.

```{r bt difference in group-specific impacts}
plot_sub_effects(bt_fit, 0.5, "BayesTree model", col_num = 5, data = x_nonbcf)
```

# BartMachine

## Tree

```{r bm tree}
bm_post <- bart_machine_get_posterior(bm_fit, as.data.frame(x_nonbcf))
bm_post$yhat <- t(bm_post$y_hat_posterior_samples)
plot_tree(bm_post, "Tree from BartMachine model", data = x_nonbcf)
```
Based on tree, we choose to investigate Z.


## Heterogeneous effects

```{r bm}
plot_het_effects(bm_post, "Comparison of estimate vs actual heterogeneous effects with bartMachine", col_num = 4, data = x_nonbcf)
```

## Differences in subgroup treatment effects

Based on the tree, we decide to look at the differences between Z smaller or greater than 0.5

```{r bm difference in group-specific impacts}
plot_sub_effects(bm_post, 0.5, "bartMachine model", 4, data = x_nonbcf)
```


# Dbarts

## Tree

```{r dbarts tree}
dbart_fit$yhat <- dbart_fit$yhat.train
plot_tree(dbart_fit, "Tree from dbarts model", data = x_nonbcf)
```
Based on tree, we choose to investigate pihat.


## Heterogeneous effects

```{r dbarts}
plot_het_effects(dbart_fit, "Comparison of estimate vs actual heterogeneous effects with dbarts", col_num = 5, data = x_nonbcf)
```

## Differences in subgroup treatment effects

Based on the tree, we decide to look at the differences between pihat smaller or greater than 0.5.

```{r dbart difference in group-specific impacts}
plot_sub_effects(dbart_fit, 0.5, "bartMachine model", 5, data = x_nonbcf)
```
