---
title: "CalCOFI kaggle challenge"
output: 
  revealjs::revealjs_presentation:
    theme: moon
    includes:
    self_contained: false
    reveal_plugins: ["notes", "menu"]
    reveal_options:
      slideNumber: true
      progress: true
      history: true
      overview: true
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)

# Libraries
library(tidyverse)
library(ggplot2)
library(corrplot)
library(gridExtra)
library(stargazer)
library(mgcv)
library(jtools)
library(lubridate)
library(ggmap)

# directories
calcofi_dir <- "C:/Users/cdelannoy/Documents/Projects/courses/bayes/kaggle/lm_calcofi/"
input_dir <- file.path(calcofi_dir, "input")
output_dir <- file.path(calcofi_dir, "output")
```

```{r loading data and preliminary processing}
calcofi_raw <- read_csv(file.path(input_dir, "bottle.csv"))
calcofi_cast<- read_csv(file.path(input_dir, "cast.csv"))

# Getting rid of "quality code" and "precision" variables, irrelevant variables (and replicate variables??) to only keep actual measures
calcofi_df <- calcofi_raw %>% 
  select(-ends_with("Aq"),
         -ends_with("1q"),
         -ends_with("2q"),
         -ends_with("3q"),
         -ends_with("4q"),
         -ends_with("Ap"),
         -ends_with("1p"),
         -ends_with("2p"),
         -ends_with("3p"),
         -ends_with("4p"),
         -ends_with("qual"),
         -ends_with("prec"),
         -ends_with("tq"),
         -ends_with("qua"),
         -ends_with("qu"),
         -`DIC Quality Comment`,
         -R_SAMP,
         -Cst_Cnt,
         -Btl_Cnt,
         -Depth_ID,
         -Sta_ID,
         -starts_with("DIC"),
         -starts_with("TA"),
         -starts_with("MeanA"),
         -IncTim,
         -starts_with("C14A") #,
         #-R_SALINITY
         ) %>% 
 mutate_if(is.character, as.numeric) %>% 
  filter(R_SALINITY > 5)

```

```{r cast cleaning}
calcofi_cast_clean <- calcofi_cast %>% 
  mutate_at(vars(Date), mdy)

```

* 60 years of oceanographic data
* Question is can we predict temperature based on salinity?
* Other variables of interest include O2 concentration and depth
* Data is non-linear and unwieldy
* Can we incorporate spatial/temporal data to fit a model? If so, how?


```{r}
outcome_na <- sum(is.na(calcofi_df$T_degC))
calcofi_nona <- calcofi_raw %>% 
  filter(!is.na(T_degC)) %>% 
  filter(R_SALINITY > 4.57)

na_table <- calcofi_df %>% 
  summarise_all(funs(sum(is.na(.))/n())) %>% 
  gather(variable, sum_na) %>% 
  filter(sum_na > 0.5)

cols_to_remove <- na_table$variable


calcofi_corrtrim <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>% 
  na.omit()

calcofi_reg <- calcofi_df[, colnames(calcofi_df[,!colnames(calcofi_df) %in% cols_to_remove])] %>%  
  select(-R_TEMP,
         -R_Depth,
         -R_O2,
         -R_PRES,
         -R_POTEMP,
         -R_SIGMA,
         -R_SVA,
         -R_DYNHT,
         -STheta,
        # -`Oxy_µmol/Kg`,
         -RecInd,
         -R_O2,
         -R_O2Sat,
         -O2Sat,
         -R_SALINITY) 

calcofi_reg <- calcofi_reg[, -5] %>% # removing `Oxy_µmol/Kg` column
  na.omit()

stopifnot(!("Oxy" %in% colnames(calcofi_reg)))
```

## Temperature vs salinity
```{r} 
ggplot(calcofi_reg, aes(y = T_degC, x = Salnty)) +
  geom_point() +
  ggtitle("Temperature vs salinity")
```

## Sections
```{r create cut-off variable}
calcofi_reg <- calcofi_reg %>% 
  mutate(sections = factor(case_when(Depthm <= 250 & O2ml_L >= 6                               ~ 1,
                                     Depthm <= 20 & O2ml_L < 6 & O2ml_L >= 5.5                 ~ 2,
                                     Depthm > 20 & Depthm <= 250 & O2ml_L < 6 & O2ml_L >= 5.5  ~ 3,
                                     Depthm <= 250 & O2ml_L < 5.5 & O2ml_L >= 5                ~ 4,
                                     Depthm <= 250 & O2ml_L < 5 & O2ml_L >= 4                  ~ 5,
                                     Depthm <= 250 & O2ml_L < 4 & O2ml_L >= 3                  ~ 6,
                                     Depthm <= 250 & O2ml_L < 3 & O2ml_L >= 2                  ~ 7,
                                     Depthm <= 250 & O2ml_L < 2 & O2ml_L >= 1.5                ~ 8,
                                     Depthm <= 250 & O2ml_L < 1.5                              ~ 9,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 6 & O2ml_L >= 2   ~ 10,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 2 & O2ml_L >= 1.5 ~ 11,
                                     Depthm > 250 & Depthm <= 500 & O2ml_L < 1.5               ~ 12,
                                     Depthm > 500 & Depthm < 1500                              ~ 13,
                                     Depthm >= 1500                                            ~ 14)))


```


```{r} 
ggplot(calcofi_reg, aes(x = Depthm, y = O2ml_L)) +
  geom_point(aes(color = sections)) +
  ggtitle("Cut-off for values of Depth for next graphs")
```


## Distribution of O2 vs Depth for each section

```{r}
ggplot(calcofi_reg, aes(y = T_degC, x = Salnty)) +
  geom_point() +
  facet_wrap(~sections, ncol = 5) +
  ggtitle("Temperature vs Salinity for various values of O2 concentration and Depth") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Time data
```{r}
samples_month <- calcofi_cast_clean %>% 
  group_by(Month) %>% 
  summarise(n_samples = n())

ggplot(samples_month, aes(x = Month, y = n_samples)) +
  geom_line() +
  ggtitle("Number of samples per month") +
  scale_x_discrete(limits = as.character(unique(samples_month$Month) %>% order())) +
  ylab("Number of samples")

```

## Location data
```{r}
mapWorld <- ggplot() +
  borders("world", colour = "gray50", fill = "gray50") +
  geom_point(data = calcofi_cast_clean, aes(x = Lon_Deg, y = Lat_Deg), colour = "red", size = 3) +
  ggtitle("Location of water samples")

mapWorld
```

## Key for what each section number represents

| Key | Depthm | O2 concentration|
|---|---|---|
| 1  | less than 250m     | greater than 6ml/L|
| 2  | less than 20m      | btw 5.5 and 6ml/L|
| 3  | btw 20 and 250m    | btw 5.5 and 6ml/L|
| 4  | less than 250m     | btw 5 and 5.5ml/L|
| 5  | less than 250m     | btw 4 and 5ml/L|
| 6  | less than 250m     | btw 3 and 4 ml/L|
| 7  | less than 250m     | btw 2 and 3 ml/L|
| 8  | less than 250m     | btw 1.5 and 2 ml/L|
| 9  | less than 250m     | less than 1.5ml/L|
| 10  | btw 250 and 500m  | btw 2 and 6 ml/L|
| 11 | btw 250 and 500m   | btw 1.5 and 2 ml/L|
| 12 | btw 250 and 500m   | less than 1.5ml/L|
| 13 | btw 500 and 1500m  | any |
| 14 | greater than 1500m | any |

## Amount of data in each section

```{r}
N_per_section <- calcofi_reg %>% 
  group_by(sections) %>% 
  summarise(sample = n())

knitr::kable(N_per_section, row.names = FALSE)
```